---
title: "Bayesian modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(BAS)
library(tidyr)
library(gridExtra)
library(MASS)
library(GGally)
```

### Load data

```{r load-data}
load("movies.Rdata")
```

* * *

## Part 1: Data
IMDb and Rotten Tomatoes are two of the most popular rating websites for movies. Both of these websites provide two sets of ratings: critics scores and audience scores, as well as numerous other variables about the movies. Our dataset is comprised of 651 randomly sampled movies produced and released before 2016. Each row in the dataset corresponds to a movie, and every column of the table represents a particular characteristic of a movie. There are 32 available variables.

* **Generalizability**: Conclusions made from this data set are generalizable because movies were randomly selected. The results are generalizable to movies produced and released before 2016.

* **Causality**: This is an observational study with no random assignment, therefore we cannot infer causality.

* **Potential biases*: Most rating systems suffer from self-selection bias since only highly motivated individuals commit to voting and ranking the movies. In addition, one person can post multiple comments and affect the ranking.

* * *

## Part 2: Data manipulation

We are first going to perform feature engineering to create new variables for our exploratory data analysis. Below is the summary of new variables, derived from existing variables in the dataset.

* *feature_film*: "yes" if title type is feature film, "no" otherwise
* *drama*: "yes" if film genre is drama, "no" otherwise runtime
* *mpaa_rating_R*: "yes" for R rated movie, "no" otherwise
* *oscar_season*: "yes" if movie is released in theaters in November, October, or December, "no" otherwise
* *summer_season*: "yes" if movie is released in theaters in May, June, July, or August, "no" otherwise

```{r}
movies <- movies %>%
  mutate(feature_film = ifelse(title_type == "Feature Film", "yes", "no"),
         drama = ifelse(genre == "Drama", "yes", "no"),
         mpaa_rating_R = ifelse(mpaa_rating == "R","yes","no"),
         oscar_season = ifelse(thtr_rel_month == 11 | thtr_rel_month == 10 | thtr_rel_month == 12, "yes", "no"),
         summer_season = ifelse(thtr_rel_month == 5 | thtr_rel_month == 6 | thtr_rel_month == 7 | thtr_rel_month == 8, "yes","no"))
```

* * *

## Part 3: Exploratory data analysis

Dataframe "movies" consists of 651 observations of 37 variables. 

```{r}
str(movies)
```

For conducting the exploratory data analysis of the relationship between audience_score and the new variables, we will create a subset of the full dataset, only containing the required data.

```{r}
df <- movies[c("feature_film", "drama", "mpaa_rating_R", "oscar_season", "summer_season", "audience_score")]
summary(df)
```

According to this summary, feature films takes the majority of title types. The drama and mpaa_rating categories are evenly represented in the dataset while the feature_film, oscar_season and summer_season seem to have a bias towards one of the "no" or "yes" bins. Audience score has the median slightly higher than mean, so the sample is probably left-skewed. The range is 11-97.

Next step is using boxplots to visualize interaction of newly-formed variables and audience_score. Boxplot visualization are used because we have numerical and categorical variables.

```{r}
p1 <- ggplot(data=movies, aes(x=movies$feature_film,y=movies$audience_score)) + geom_boxplot()
p2 <- ggplot(data=movies, aes(x=movies$drama,y=movies$audience_score)) + geom_boxplot()
p3 <- ggplot(data=movies, aes(x=movies$mpaa_rating_R,y=movies$audience_score)) + geom_boxplot()
p4 <- ggplot(data=movies, aes(x=movies$oscar_season,y=movies$audience_score)) + geom_boxplot()
p5 <- ggplot(data=movies, aes(x=movies$summer_season,y=movies$audience_score)) + geom_boxplot()

grid.arrange(p1, p2, p3, p4, p5, ncol=3)
```

* This boxplot shows that apart from a few outliers, movies that are not feature films tend to garner a bigger audience score. The feature movies group have a lower audience rating.

* Drama movies seem to have higher audience rating and a bigger median audience score with similar variability.

* R rated movies seem to have a lower median audience score but the results are mostly similar. There is no difference in the audience score between whether a movie is R rated or not.

* Movies in the Oscar season have a higher audience score with similar variability to those who are not in the Oscar season. 

* It seems there is no difference in the audience score between whether a movie is in the summer season or not. Movies in summer season appear to have a smaller median audience score than those not in summer season.

* * *

## Part 4: Modeling

Next step will be developing a Bayesian Regression Model to predict the audience_score from the variables mentioned above. This model is created as follows:

```{r}
moviesNan = na.omit(movies)
modelReg = lm( data=moviesNan, log(audience_score) ~ feature_film + drama + runtime
                  + mpaa_rating_R + thtr_rel_year + oscar_season + summer_season
                  + imdb_rating + imdb_num_votes + critics_score + best_pic_nom
                  + best_pic_win + best_actor_win + best_actress_win 
                  + best_dir_win + top200_box)
BIC(modelReg)
summary(modelReg)
```

The BIC of the full model is -117.7922.

Next step is backward elimination with Bayesian Information Criterion (BIC). Every time we will eliminate the variable that reduces the BIC the most until no variables can decrease BIC.

```{r}
elimination <- stepAIC(modelReg, trace=0)
elimination
```

Based on backward elimination with BIC, explanatory variables included in the final model are feature_film, runtime, mpaa_rating_R, imdb_rating, and critics_score.

```{r}
ggplot(data = elimination, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")

qqnorm(elimination$residuals)
qqline(elimination$residuals)
```

According to the diagnostics, there are residuals with large values for low BMA predictions, indicated that the observed values are greater than the predicted values. The assumption of equal variance also seems to not being fulfilled in this model due to the residuals not being spread equally along the range of the predictors, signifying the presence of a non-linear relationship between the predictors.

```{r}
elimination$coefficients
```

If the movie is feature, then the expected log of the audience score will increase by 0.06 while keeping all other variables the same. If the runtime increase by 1 minutes, then the expected log of the audience score will decrease by -0.0013 while keeping all other variables the same.

* * *

## Part 5: Prediction

We are going to use the model developed earlier to predict the audience score for the new movie from 2016 that is not included in the sample. Movie "Moonlight", categorized as drama, on the Rotten Tomatoes website has a 98% rating (critics score). New data frame will be created for this movie.

```{r}
moviePredict <- data.frame(runtime = 110, mpaa_rating_R = 'yes', feature_film = 'yes', drama = 'yes', oscar_season = 'yes', summer_season = 'no', imdb_num_votes = 190729, thtr_rel_year = 2016, imdb_rating = 7.4, critics_score = 98, best_pic_nom = 'yes', best_pic_win = 'yes', best_actor_win = 'no', best_actress_win = 'no', best_dir_win = 'no', top200_box = 'no')

exp(predict(elimination, moviePredict, interval = "prediction", level = 0.95)) 
```

The model predicts, with 95% confidence, that drama "Moonlight" is expected to have an audience score between 52.88 and 117.54. Our prediction interval contains actual audience score from the Rotten Tomatoes website (79).

* * *

## Part 6: Conclusion

In the scope of this project Bayesian regression model for predicting the audience score on Rotten Tomatoes was developed. Model performed well with selected movie, resulting with the true audience score within the credible interval. Potential shortcoming is that our model's predictive power is limited because the sample data is not representative. Further, testing on only a single movie does not mean that this model performs well overall. Larger number of observations would assure better predictive value.